Universal algebra and the \texorpdfstring{$ \lambda $}{lambda}-calculus

### 2024-05-28 report/src/alternative-definitions.tex
Improve citations, quotes, move dependencies and definitions to style file

\section{Lawvere theory}
\begin{definition}
  An \iindex{algebraic theory} as presented in \cite{algebaric-theories-2010} is a small category with finite products.
\end{definition}
\begin{definition}
  An \iindex{algebra} for an algebraic theory $ T $ is a finite-product-preserving functor $ T \to \SET $.
\end{definition}

@@

\begin{definition}
  A \iindex{cartesian operad} T is a functor $ T: F \to \SET $, together with an `identity' element $ \id{}: T(1) $ and for all $ m, n_1, \dots, n_m: \mathbb N $, a composition map
  \[ T(m) \times \prod_i T(n_i) \to T\left(\sum_i n_i\right), \]
  written $ (f, g_1, \dots, g_m) \mapsto f[g_1, \dots, g_m] $, satisfying some identity, associativity and naturality conditions (see \cite{lambda-monoid}, Definition 2.1, for more details).
\end{definition}

@@

\begin{remark}
  Note that for an adjunction $ F \dashv G $, $ F \bullet G $ gives a monad. In the same way, there exists a notion of \iindex{relative adjunction}, from which we can obtain a relative monad (see \cite{monads-endofunctors}, Theorem 2.10).
\end{remark}

@@

\section{Monoid in a skew-monoidal category}
For details and a general treatment, see \cite{monads-endofunctors}, Section 3 and specifically Theorem 3.4.

@@

\begin{definition}
  Given a (skew-)monoidal category $ (C, \otimes, I) $, a \iindex{monoid} in this category is an object $ T : C $, together with a `multiplication' $ \mu: C(T \otimes T, T) $ and a `unit' morphism $ \eta: C(I, T) $ satisfying a couple of laws (see \cite{MacLane}, Section III.6).
\end{definition}

>>>>

\section{Lawvere theory}
\begin{definition}
  An \iindex{algebraic theory} as presented in \autocite{algebaric-theories-2010} is a small category with finite products.
\end{definition}
\begin{definition}
  An \iindex{algebra} for an algebraic theory $ T $ is a finite-product-preserving functor $ T \to \SET $.
\end{definition}

@@

\begin{definition}
  A \iindex{cartesian operad} T is a functor $ T: F \to \SET $, together with an `identity' element $ \id{}: T(1) $ and for all $ m, n_1, \dots, n_m: \mathbb N $, a composition map
  \[ T(m) \times \prod_i T(n_i) \to T\left(\sum_i n_i\right), \]
  written $ (f, g_1, \dots, g_m) \mapsto f[g_1, \dots, g_m] $, satisfying some identity, associativity and naturality conditions (see \autocite{lambda-monoid}, Definition 2.1, for more details).
\end{definition}

@@

\begin{remark}
  Note that for an adjunction $ F \dashv G $, $ F \bullet G $ gives a monad. In the same way, there exists a notion of \iindex{relative adjunction}, from which we can obtain a relative monad (see \autocite{monads-endofunctors}, Theorem 2.10).
\end{remark}

@@

\section{Monoid in a skew-monoidal category}
For details and a general treatment, see \autocite{monads-endofunctors}, Section 3 and specifically Theorem 3.4.

@@

\begin{definition}
  Given a (skew-)monoidal category $ (C, \otimes, I) $, a \iindex{monoid} in this category is an object $ T : C $, together with a `multiplication' $ \mu: C(T \otimes T, T) $ and a `unit' morphism $ \eta: C(I, T) $ satisfying a couple of laws (see \autocite{MacLane}, Section III.6).
\end{definition}

### 2024-05-28 report/src/category-theory.tex
Improve citations, quotes, move dependencies and definitions to style file

I will assume a familiarity with the category-theoretical concepts presented in \textcite{CT4P}. These include categories, functors, isomorphisms, natural transformations, adjunctions, equivalences and limits.

@@

One concept in category theory that can be used to describe a lot of limits and adjunctions is that of a universal arrow (see for example \textcite{MacLane}, Part III)
\begin{definition}
  A \iindex{universal arrow} from an object $ c: C $ to a functor $ F: D \to C $ consists of an object $ d: D $ and a morphism $ f: D(c, F(d)) $ such that for every similar pair $ (d^\prime, f^\prime) $, $ f^\prime $ factors uniquely as $ f \cdot F(g) $ for some $ g: C(d, d^\prime) $:
  \begin{center}
    \begin{tikzcd}
      c \arrow[d, "f"'] \arrow[rd, "f^\prime"] &\\
      F(d) \arrow[r, "F(g)"', dashed] & F(d^\prime)
    \end{tikzcd}
  \end{center}
\end{definition}

@@

An alternative characterization (see \textcite{MacLane}, chapter IV.1, Theorem 2) of an adjunction $ L \dashv R $ is as a natural bijection
\[ \varphi: D(L(c), d) \xrightarrow{\sim} C(c, R(d)). \]
Naturality means that for all $ f: C(c^\prime, c) $, $ g: D(d, d^\prime) $ and $ h: D(L(c), d) $,
\[ \varphi(L(f) \cdot h \cdot g) = f \cdot \varphi(h) \cdot R(g). \]

Lastly, one can construct an adjunction using universal arrows. This lends itself particularly well for a formalization, where it is often preferable to have as little `demonstranda' as possible:
\begin{lemma}
  One can construct an adjunction $ (L, R, \eta, \epsilon) $ as above from only the functor $ L: C \to D $ and, for each $ c: C $, a universal arrow $ (R(c), \epsilon_c) $ from $ L $ to $ c $.
\end{lemma}
\begin{proof}
  See \textcite{MacLane}, Chapter IV.1, Theorem 2 (iv).
\end{proof}

@@

\section{Yoneda}
We can embed a category $ C $ fully faithfully into the functor category $ P C = [\op C, \SET] $ as follows (see \textcite{Kashiwara}, Section 1.4):
\begin{definition}
  The \iindex{Yoneda embedding} $ Y : C \hookrightarrow P C $ is given on objects by $ Y(c) = C(-, c) $:
  \[ Y(c)(d) = C(d, c) \quad \text{and} \quad Y(c)(f)(g) = f \cdot g \]
  for $ d: C $, $ f: C(d, d^\prime) $ and $ g: C(d^\prime, c) $. It sends a morphism $ f: C(c, c^\prime) $ to the natural transformation $ Y(f): C(-, c) \Rightarrow C(-, c^\prime) $ given by
  \[ Y(f)(d)(g) = g \cdot f \]
  for $ d: C $ and $ g: C(d, c) $.
\end{definition}

Now, this embedding has a couple of properties:
\begin{lemma}
  For any $ c: C $ and $ F : PC $, we have an equivalence $ PC(Y(c), F) \simeq F(c) $, and this equivalence is natural in $ c $ and $ F $.
\end{lemma}
\begin{proof}
  See \textcite{Kashiwara}, Proposition 1.4.3.
\end{proof}

\begin{lemma}
  The Yoneda embedding functor preserves limits.
\end{lemma}
\begin{proof}
  See \textcite{borceux}, Volume 1, Proposition 2.15.5.
\end{proof}

\begin{lemma}
  The Yoneda embedding functor preserves exponential objects (\textcite{stackexchange:yoneda-exponentials}).
\end{lemma}
\begin{proof}

@@

\begin{definition}
  A morphism $ f: E(y, z) $ is called \iindex{cartesian} if for all $ g: E(x, z) $ and $ h: B(P(x), P(y)) $ with $ h \cdot P(f) = P(g) $, there exists $ \bar h: E(x, y) $ such that $ P(\bar h) = h $ and $ \bar h \cdot f = g $, like illustrated in the following diagram from \textcite{nlab:grothendieck_fibration}
  \begin{center}
    \begin{tikzcd}[sep=large]
      E \arrow[d, "P"] &x \arrow[rr, "\forall g", bend left] \arrow[r, "\exists! \bar h"', dashed] & y \arrow[r, "f"'] & z\\
      B & P(x) \arrow[r, "\forall h"'] \arrow[rr, "P(g)", bend left] & P(y) \arrow[r, "P(f)"'] & P(z)
    \end{tikzcd}
  \end{center}
\end{definition}

@@

\begin{lemma}\label{lem:delta-exponentials}
  $ \Delta_X $ preserves exponential objects:
\end{lemma}
\begin{proof}
  See \textcite{borceux}, Volume 3, Lemma 5.8.2. This lemma shows this via the following equivalences:
  \begin{align*}
    (C \downarrow X)((A, f), \Delta_X Y^Z) &\cong C(A, Y^Z)\\
    &\cong C(A \times Z, Y)\\
    &\cong (C \downarrow X)((A \times X \times Z, \langle f, p_1 \rangle), \Delta_X Y)\\
    &\cong (C \downarrow X)((A, f) \times \Delta_X Z, \Delta_X Y)\\
  \end{align*}
\end{proof}

\section{Dependent products and sums}\label{sec:dependent-products}
The following is based loosely on Section 4.1 of \textcite{taylor}.

@@

Apart from having dependent sums and products, there also is the following theorem that shows the significance of locally cartesian closedness:
\begin{lemma}\label{lem:locally-cartesian-closed}
  A category $ C $ is locally cartesian closed iff $ (C \downarrow A) $ is cartesian closed for each $ A : C $.
\end{lemma}
\begin{proof}
  See the end of Section 1.3 of \textcite{freyd}.
\end{proof}

@@

  By Proposition 1.34 in \textcite{freyd}, $ \prod_{p_1} \Delta_{X \times Z} W $ is given as the following pullback:
  \begin{center}
    \begin{tikzcd}
      \prod_{p_1} \Delta_{X \times Z} W \arrow[r] \arrow[d] & (\Delta_X Z \times W)^{\Delta_X Z} \arrow[d, "(\Delta_X p_1)^{\Delta_X Z}"]\\
      X \arrow[r] & (\Delta_X Z)^{\Delta_X Z}
    \end{tikzcd}
  \end{center}

@@

\section{Kan Extensions}
One of the most general and abstract concepts in category theory is the concept of \textit{Kan extensions}. In \textcite{MacLane}, Section X.7, MacLane notes that
\begin{quote}
  The notion of Kan extensions subsumes all the other fundamental concepts of category theory.
\end{quote}
In this thesis, we will use left Kan extension a handful of times. It comes in handy when we want to extend a functor along another functor in the following way:

@@

If a category has limits (resp. colimits), we can construct the right (resp. left) Kan extension in a `pointwise' fashion (see Theorem X.3.1 in \textcite{MacLane} or Theorem 2.3.3 in \textcite{Kashiwara}). Below, I will outline the parts of the construction that we will need explicitly in this thesis.
\begin{lemma}
  If $ C $ has colimits, $ \Lan F {} $ exists.
\end{lemma}
\begin{proof}
  First of all, for objects $ b: B $, we take
  \[ \Lan F G(b) := \text{colim} \left( (F \downarrow b) \to A \xrightarrow G C \right). \]

@@

\begin{lemma}\label{lem:lan-precomp-iso}
  If $ F : A \to B $ is a fully faithful functor, and $ C $ is a category with colimits, $ \eta: \id{[A, C]} \Rightarrow \Lan F {} \bullet F_* $ is a natural isomorphism.
\end{lemma}
\begin{proof}
  To show that $ \eta $ is a natural isomorphism, we have to show that $ \eta_G(a^\prime): G(a^\prime) \Rightarrow \Lan F G(F(a^\prime)) $ is an isomorphism for all $ G: [A, C] $ and $ a^\prime: A $. Since a left adjoint is unique up to natural isomorphism (\textcite{CT4P}, Exercise 153), we can assume that $ \Lan F G(F(a^\prime)) $ is given by
  \[ \text{colim} ((F \downarrow F(a^\prime)) \to A \xrightarrow G C). \]
  Now, the diagram for this colimit consists of $ G(a) $ for each arrow $ f: B(F(a), F(a^\prime)) $. Since $ F $ is fully faithful, we have $ f = F(\overline f) $ for some $ \overline f: A(a, a^\prime) $. If we now take the arrows $ G(\overline f): C(G(a), G(a^\prime)) $, the universal property of the colimit gives an arrow
  \[ \varphi: C(\Lan F G(F(a^\prime)), G(a^\prime)) \]
  which constitutes an inverse to $ \eta_G(a^\prime) $. The proof of this revolves around properties of the colimit and its (induced) morphisms.
\end{proof}

@@

\begin{lemma}\label{lem:retracts-rezk}
  $ \hat C $ is the Rezk completion of $ \overline C $.
\end{lemma}
\begin{proof}
  First of all, note that $ \hat C $ is univalent, since it is a presheaf category (see Theorem 4.5 in \textcite{univalent-categories}). So we must show that we have a weak equivalence from $ \overline C $ to $ \hat C $.

>>>>

I will assume a familiarity with the category-theoretical concepts presented in \autocite{CT4P}. These include categories, functors, isomorphisms, natural transformations, adjunctions, equivalences and limits.

@@

One concept in category theory that can be used to describe a lot of limits and adjunctions is that of a universal arrow (see for example \autocite{MacLane}, Part III)
\begin{definition}
  A \iindex{universal arrow} from an object $ c: C $ to a functor $ F: D \to C $ consists of an object $ d: D $ and a morphism $ f: D(c, F(d)) $ such that for every similar pair $ (d^\prime, f^\prime) $, $ f^\prime $ factors uniquely as $ f \cdot F(g) $ for some $ g: C(d, d^\prime) $:
  \begin{center}
    \begin{tikzcd}
      c \arrow[d, "f"'] \arrow[rd, "f^\prime"] &\\
      F(d) \arrow[r, "F(g)"', dashed] & F(d^\prime)
    \end{tikzcd}
  \end{center}
\end{definition}

@@

An alternative characterization (see \autocite{MacLane}, chapter IV.1, Theorem 2) of an adjunction $ L \dashv R $ is as a natural bijection
\[ \varphi: D(L(c), d) \xrightarrow{\sim} C(c, R(d)). \]
Naturality means that for all $ f: C(c^\prime, c) $, $ g: D(d, d^\prime) $ and $ h: D(L(c), d) $,
\[ \varphi(L(f) \cdot h \cdot g) = f \cdot \varphi(h) \cdot R(g). \]

Lastly, one can construct an adjunction using universal arrows. This lends itself particularly well for a formalization, where it is often preferable to have as little `demonstranda' as possible:
\begin{lemma}
  One can construct an adjunction $ (L, R, \eta, \epsilon) $ as above from only the functor $ L: C \to D $ and, for each $ c: C $, a universal arrow $ (R(c), \epsilon_c) $ from $ L $ to $ c $.
\end{lemma}
\begin{proof}
  See \autocite{MacLane}, Chapter IV.1, Theorem 2 (iv).
\end{proof}

@@

\section{Yoneda}
We can embed a category $ C $ fully faithfully into the functor category $ P C = [\op C, \SET] $ as follows (see \autocite{Kashiwara}, Section 1.4):
\begin{definition}
  The \iindex{Yoneda embedding} $ Y : C \hookrightarrow P C $ is given on objects by $ Y(c) = C(-, c) $:
  \[ Y(c)(d) = C(d, c) \quad \text{and} \quad Y(c)(f)(g) = f \cdot g \]
  for $ d: C $, $ f: C(d, d^\prime) $ and $ g: C(d^\prime, c) $. It sends a morphism $ f: C(c, c^\prime) $ to the natural transformation $ Y(f): C(-, c) \Rightarrow C(-, c^\prime) $ given by
  \[ Y(f)(d)(g) = g \cdot f \]
  for $ d: C $ and $ g: C(d, c) $.
\end{definition}

Now, this embedding has a couple of properties:
\begin{lemma}
  For any $ c: C $ and $ F : PC $, we have an equivalence $ PC(Y(c), F) \simeq F(c) $, and this equivalence is natural in $ c $ and $ F $.
\end{lemma}
\begin{proof}
  See \autocite{Kashiwara}, Proposition 1.4.3.
\end{proof}

\begin{lemma}
  The Yoneda embedding functor preserves limits.
\end{lemma}
\begin{proof}
  See \autocite{borceux}, Volume 1, Proposition 2.15.5.
\end{proof}

\begin{lemma}
  The Yoneda embedding functor preserves exponential objects (\autocite{stackexchange:yoneda-exponentials}).
\end{lemma}
\begin{proof}

@@

\begin{definition}
  A morphism $ f: E(y, z) $ is called \iindex{cartesian} if for all $ g: E(x, z) $ and $ h: B(P(x), P(y)) $ with $ h \cdot P(f) = P(g) $, there exists $ \bar h: E(x, y) $ such that $ P(\bar h) = h $ and $ \bar h \cdot f = g $, like illustrated in the following diagram from \autocite{nlab:grothendieck_fibration}
  \begin{center}
    \begin{tikzcd}[sep=large]
      E \arrow[d, "P"] &x \arrow[rr, "\forall g", bend left] \arrow[r, "\exists! \bar h"', dashed] & y \arrow[r, "f"'] & z\\
      B & P(x) \arrow[r, "\forall h"'] \arrow[rr, "P(g)", bend left] & P(y) \arrow[r, "P(f)"'] & P(z)
    \end{tikzcd}
  \end{center}
\end{definition}

@@

\begin{lemma}\label{lem:delta-exponentials}
  $ \Delta_X $ preserves exponential objects:
\end{lemma}
\begin{proof}
  See \autocite{borceux}, Volume 3, Lemma 5.8.2. This lemma shows this via the following equivalences:
  \begin{align*}
    (C \downarrow X)((A, f), \Delta_X Y^Z) &\cong C(A, Y^Z)\\
    &\cong C(A \times Z, Y)\\
    &\cong (C \downarrow X)((A \times X \times Z, \langle f, p_1 \rangle), \Delta_X Y)\\
    &\cong (C \downarrow X)((A, f) \times \Delta_X Z, \Delta_X Y)\\
  \end{align*}
\end{proof}

\section{Dependent products and sums}\label{sec:dependent-products}
The following is based loosely on Section 4.1 of \autocite{taylor}.

@@

Apart from having dependent sums and products, there also is the following theorem that shows the significance of locally cartesian closedness:
\begin{lemma}\label{lem:locally-cartesian-closed}
  A category $ C $ is locally cartesian closed iff $ (C \downarrow A) $ is cartesian closed for each $ A : C $.
\end{lemma}
\begin{proof}
  See the end of Section 1.3 of \autocite{freyd}.
\end{proof}

@@

  By Proposition 1.34 in \autocite{freyd}, $ \prod_{p_1} \Delta_{X \times Z} W $ is given as the following pullback:
  \begin{center}
    \begin{tikzcd}
      \prod_{p_1} \Delta_{X \times Z} W \arrow[r] \arrow[d] & (\Delta_X Z \times W)^{\Delta_X Z} \arrow[d, "(\Delta_X p_1)^{\Delta_X Z}"]\\
      X \arrow[r] & (\Delta_X Z)^{\Delta_X Z}
    \end{tikzcd}
  \end{center}

@@

\section{Kan Extensions}
One of the most general and abstract concepts in category theory is the concept of \textit{Kan extensions}. In \autocite{MacLane}, Section X.7, MacLane notes that

\enquote{The notion of Kan extensions subsumes all the other fundamental concepts of category theory.}

In this thesis, we will use left Kan extension a handful of times. It comes in handy when we want to extend a functor along another functor in the following way:

@@

If a category has limits (resp. colimits), we can construct the right (resp. left) Kan extension in a `pointwise' fashion (see Theorem X.3.1 in \autocite{MacLane} or Theorem 2.3.3 in \autocite{Kashiwara}). Below, I will outline the parts of the construction that we will need explicitly in this thesis.
\begin{lemma}
  If $ C $ has colimits, $ \Lan F {} $ exists.
\end{lemma}
\begin{proof}
  First of all, for objects $ b: B $, we take
  \[ \Lan F G(b) := \text{colim} \left( (F \downarrow b) \to A \xrightarrow G C \right). \]

@@

\begin{lemma}\label{lem:lan-precomp-iso}
  If $ F : A \to B $ is a fully faithful functor, and $ C $ is a category with colimits, $ \eta: \id{[A, C]} \Rightarrow \Lan F {} \bullet F_* $ is a natural isomorphism.
\end{lemma}
\begin{proof}
  To show that $ \eta $ is a natural isomorphism, we have to show that $ \eta_G(a^\prime): G(a^\prime) \Rightarrow \Lan F G(F(a^\prime)) $ is an isomorphism for all $ G: [A, C] $ and $ a^\prime: A $. Since a left adjoint is unique up to natural isomorphism (\autocite{CT4P}, Exercise 153), we can assume that $ \Lan F G(F(a^\prime)) $ is given by
  \[ \text{colim} ((F \downarrow F(a^\prime)) \to A \xrightarrow G C). \]
  Now, the diagram for this colimit consists of $ G(a) $ for each arrow $ f: B(F(a), F(a^\prime)) $. Since $ F $ is fully faithful, we have $ f = F(\overline f) $ for some $ \overline f: A(a, a^\prime) $. If we now take the arrows $ G(\overline f): C(G(a), G(a^\prime)) $, the universal property of the colimit gives an arrow
  \[ \varphi: C(\Lan F G(F(a^\prime)), G(a^\prime)) \]
  which constitutes an inverse to $ \eta_G(a^\prime) $. The proof of this revolves around properties of the colimit and its (induced) morphisms.
\end{proof}

@@

\begin{lemma}\label{lem:retracts-rezk}
  $ \hat C $ is the Rezk completion of $ \overline C $.
\end{lemma}
\begin{proof}
  First of all, note that $ \hat C $ is univalent, since it is a presheaf category (see Theorem 4.5 in \autocite{univalent-categories}). So we must show that we have a weak equivalence from $ \overline C $ to $ \hat C $.

### 2024-05-28 report/src/paper.tex
Improve citations, quotes, move dependencies and definitions to style file

Recall that $ \Pshf L $ is isomorphic to the presheaf category $ P \mathbf L $, so it is an elementary topos, so it is locally cartesian closed (Example 5.2.5 and Theorem 5.8.4, \textcite{borceux}, Volume 3). Therefore, for a morphism $ f: \Pshf L(Y, X) $, we have adjunctions
\begin{center}
  \begin{tikzcd}
    (\Pshf L \downarrow X) \arrow[r, "f^*"' description, ""{name=B}] & (\Pshf L \downarrow Y) \arrow[l, bend right, "{\sum_f = - \cdot f}"'{name=A}] \arrow[l, bend left, "\prod_f"{name=C}]
    \arrow[from=A, to=B, symbol=\dashv]
    \arrow[from=B, to=C, symbol=\dashv]
  \end{tikzcd}
\end{center}

@@

  However, as mentioned in \textcite{univalent-categories} halfway through the introduction, and as explained more thoroughly in \textcite{hottbook}, Lemma 3.8.5, if $ \R(\varphi(X)) $ is not a set (and if $ L $ is nontrivial, it is never a set), then there is no axiom of choice which can turn the `for all \dots, there exists \dots' into a `there exists a function'.

@@

  However, showing that it is a $ \lambda $-theory morphism is a different matter. Hyland claims
  \begin{quote}
    $ F $ preserves $ \mathbf 1_n $, which determines the function space as a retract of the universal. So $ \bar F $ preserves the retract and the result follows.
  \end{quote}
  Although this covers the core of the argument, it is very complicated to actually verify that this works, because we need to pass through a lot of isomorphisms, and check that they work nicely together:
  \begin{align*}
    \alpha_A: \RAct{A_1}(X \times Y, Z) &\xrightarrow{\sim} \RAct{A_1}(X, Z^Y);\\
    \beta: F_*(U_A) &\xrightarrow{\sim} U_{A^\prime};\\
    \bar \gamma: F_*(A \times B) &\xrightarrow{\sim} F_*(A) \times F_*(B);\\
    \gamma_n: F_*(X^n) &\xrightarrow{\sim} F_*(X)^n;\\
    % \delta: F_*(X^Y) &\xrightarrow{\sim} F_*(X)^{F_*(Y)};\\
    \delta_A: U_A^{U_A} &\xrightarrow{\sim} A_2.
  \end{align*}
  Since we already motivated this in a different way, we will leave it as an exercise for the enthousiastic reader.
  % with $ \gamma_{n + 1} = \bar{\gamma} \cdot (\gamma_n \times \id X) $.
  % Then, for $ s: E(U_A)_{n + 1} $, we have $ \lambda_A(s) = \alpha_A(s) \cdot \delta_A \cdot \pi_A : \RAct{A_1}(U_A^n, U_A) $ for the projection $ \pi_A: A_2 \to U_A $. However, note that $ f^*(\lambda_A(s)): \RAct{A^\prime_1}(f^*(U_{A^\prime}^n), f^*(U_{A^\prime})) $. To compare this to $ \lambda_{A^\prime}(\dots) $, we actually need to take
  % \[ (\beta^{-1})^n \cdot \gamma_n^{-1} \cdot f^*(\lambda_A(s)) \cdot \beta : \RAct{A^\prime_1}(U_A^n, U_A). \]
  % On the other hand, we also have $ f^*(s): \RAct{A^\prime_1}(f^*(U_A^{n + 1}), f^*(U_A)) $ and we cannot apply $ \lambda_{A^\prime} $ directly. Instead, the term that we end up with is
  % \[ \lambda_{A^\prime}((\beta^{-1})^{n + 1} \cdot \gamma_{n + 1}^{-1} \cdot f^*(s) \cdot \beta): \RAct{A^\prime_1}(U_A^n, U_A). \]
  % So the equality that we need to prove is (using functoriality of $ f^* $ and naturality of $ \alpha_{A^\prime} $):
  % \begin{align*}
  %   &(\beta^{-1})^n \cdot \gamma_n^{-1} \cdot f^*(\alpha_A(s)) \cdot f^*(\delta_A \cdot \pi_A) \cdot \beta\\
  %   &= (\beta^{-1})^n \cdot \gamma_n^{-1} \cdot \alpha_{A^\prime}((\id X^n \times \beta^{-1}) \cdot \bar{\gamma}^{-1} \cdot f^*(s)) \cdot \beta^{U_A} \cdot \delta_{A^\prime} \cdot \pi_{A^\prime}.
  % \end{align*}
  % Because of the complicated definition of each of these terms, I was unable to verify the correctness of this statement (and the similar statement about $ \rho: E(U_A)_n \to E(U_A)_{n + 1} $) within a day, even though I am willing to believe Hyland's claim.

@@

By Lemma 3.2 in \textcite{nlab:adjoint_equivalence} and Lemma \ref{lem:zigzag}, $ \eta $ and $ \epsilon $ satisfy both zigazg identities, and we can state the fundamental theorem of the $ \lambda $-calculus:
\begin{theorem}
  There is an adjoint equivalence $ \LamTh \cong \Alg{\Lambda} $, sending a $ \lambda $-theory $ L $ to the $ \Lambda $-algebra $ L_0 $, with an inverse functor that sends a $ \Lambda $-algebra $ A $ to the theory $ E_{\RAct{A_1}}(U_A) $.
\end{theorem}

>>>>

Recall that $ \Pshf L $ is isomorphic to the presheaf category $ P \mathbf L $, so it is an elementary topos, so it is locally cartesian closed (Example 5.2.5 and Theorem 5.8.4, \autocite{borceux}, Volume 3). Therefore, for a morphism $ f: \Pshf L(Y, X) $, we have adjunctions
\begin{center}
  \begin{tikzcd}
    (\Pshf L \downarrow X) \arrow[r, "f^*"' description, ""{name=B}] & (\Pshf L \downarrow Y) \arrow[l, bend right, "{\sum_f = - \cdot f}"'{name=A}] \arrow[l, bend left, "\prod_f"{name=C}]
    \arrow[from=A, to=B, symbol=\dashv]
    \arrow[from=B, to=C, symbol=\dashv]
  \end{tikzcd}
\end{center}

@@

  However, as mentioned in \autocite{univalent-categories} halfway through the introduction, and as explained more thoroughly in \autocite{hottbook}, Lemma 3.8.5, if $ \R(\varphi(X)) $ is not a set (and if $ L $ is nontrivial, it is never a set), then there is no axiom of choice which can turn the `for all \dots, there exists \dots' into a `there exists a function'.

@@

  However, showing that it is a $ \lambda $-theory morphism is a different matter. Hyland claims

  \enquote{$ F $ preserves $ \mathbf 1_n $, which determines the function space as a retract of the universal. So $ \bar F $ preserves the retract and the result follows.}

  Although this covers the core of the argument, it is very complicated to actually verify that this works, because we need to pass through a lot of isomorphisms, and check that they work nicely together:
  \begin{align*}
    \alpha_A: \RAct{A_1}(X \times Y, Z) &\xrightarrow{\sim} \RAct{A_1}(X, Z^Y);\\
    \beta: F_*(U_A) &\xrightarrow{\sim} U_{A^\prime};\\
    \bar \gamma: F_*(A \times B) &\xrightarrow{\sim} F_*(A) \times F_*(B);\\
    \gamma_n: F_*(X^n) &\xrightarrow{\sim} F_*(X)^n;\\
    % \delta: F_*(X^Y) &\xrightarrow{\sim} F_*(X)^{F_*(Y)};\\
    \delta_A: U_A^{U_A} &\xrightarrow{\sim} A_2.
  \end{align*}
  Since we already motivated this in a different way, we will leave it as an exercise for the enthousiastic reader.
  % with $ \gamma_{n + 1} = \bar{\gamma} \cdot (\gamma_n \times \id X) $.
  % Then, for $ s: E(U_A)_{n + 1} $, we have $ \lambda_A(s) = \alpha_A(s) \cdot \delta_A \cdot \pi_A : \RAct{A_1}(U_A^n, U_A) $ for the projection $ \pi_A: A_2 \to U_A $. However, note that $ f^*(\lambda_A(s)): \RAct{A^\prime_1}(f^*(U_{A^\prime}^n), f^*(U_{A^\prime})) $. To compare this to $ \lambda_{A^\prime}(\dots) $, we actually need to take
  % \[ (\beta^{-1})^n \cdot \gamma_n^{-1} \cdot f^*(\lambda_A(s)) \cdot \beta : \RAct{A^\prime_1}(U_A^n, U_A). \]
  % On the other hand, we also have $ f^*(s): \RAct{A^\prime_1}(f^*(U_A^{n + 1}), f^*(U_A)) $ and we cannot apply $ \lambda_{A^\prime} $ directly. Instead, the term that we end up with is
  % \[ \lambda_{A^\prime}((\beta^{-1})^{n + 1} \cdot \gamma_{n + 1}^{-1} \cdot f^*(s) \cdot \beta): \RAct{A^\prime_1}(U_A^n, U_A). \]
  % So the equality that we need to prove is (using functoriality of $ f^* $ and naturality of $ \alpha_{A^\prime} $):
  % \begin{align*}
  %   &(\beta^{-1})^n \cdot \gamma_n^{-1} \cdot f^*(\alpha_A(s)) \cdot f^*(\delta_A \cdot \pi_A) \cdot \beta\\
  %   &= (\beta^{-1})^n \cdot \gamma_n^{-1} \cdot \alpha_{A^\prime}((\id X^n \times \beta^{-1}) \cdot \bar{\gamma}^{-1} \cdot f^*(s)) \cdot \beta^{U_A} \cdot \delta_{A^\prime} \cdot \pi_{A^\prime}.
  % \end{align*}
  % Because of the complicated definition of each of these terms, I was unable to verify the correctness of this statement (and the similar statement about $ \rho: E(U_A)_n \to E(U_A)_{n + 1} $) within a day, even though I am willing to believe Hyland's claim.

@@

By Lemma 3.2 in \autocite{nlab:adjoint_equivalence} and Lemma \ref{lem:zigzag}, $ \eta $ and $ \epsilon $ satisfy both zigazg identities, and we can state the fundamental theorem of the $ \lambda $-calculus:
\begin{theorem}
  There is an adjoint equivalence $ \LamTh \cong \Alg{\Lambda} $, sending a $ \lambda $-theory $ L $ to the $ \Lambda $-algebra $ L_0 $, with an inverse functor that sends a $ \Lambda $-algebra $ A $ to the theory $ E_{\RAct{A_1}}(U_A) $.
\end{theorem}
